Storm surge modelling in the Strait of Georgia

### 2014-07-24 RoughWork/manuscript.tex
Excluded the background state from tide-surge comparison

\subsection{Model domain}
The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells. As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.

@@

\subsection{Boundary conditions and sub-grid scale processes}
The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$), temperature and salinity climatologies, and the sea surface height anomaly. Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M$_2$ and K$_1$ constituents. Additionally, O$_1$ and S$_2$ elevation harmonics from their measurements were employed. The remaining constituents were extrapolated from Webtide. 
% Needs editing! Update when we are finished with tides.

@@

Several significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure. These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events. In addition, the contribution to the surge of several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed. We have also examined the spatial variabilty of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.

@@

Finally, an assessment of the importance of including tide-surge interactions is presented next. A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}. Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only. A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction. This figure shows there is almost no difference between the simulations at all locations, rendering the tide-sure interaction negligible. In operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction. 
% I'm not sure if removing a background state is really appropriate here. There could be some differences in the background states because of the tides(ie where fresh water moves to, etc). I also think it is more appropriate to use the atmospheric pressure to remove the background state...

%Finally, an assessment of the importance of including tide-surge interactions is presented next. A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}. Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only. A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction. The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm higher than the maximum of the surge only case.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the modelled residual and surge only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation. However, the effect is a relatively small underestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006_tidesurge_bg.pdf}
\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed).}
\label{fig:tidesurge}
\end{figure}

>>>>

The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells. As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.

@@

The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$), temperature and salinity climatologies, and the sea surface height anomaly. Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M$_2$ and K$_1$ constituents. Additionally, O$_1$ and S$_2$ elevation harmonics from their measurements were employed. The remaining constituents were extrapolated from Webtide. 
% Needs editing! Update when we are finished with tides.

@@

Several significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure. These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events. In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed. We have also examined the spatial variabilty of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.

@@

%Finally, an assessment of the importance of including tide-surge interactions is presented next. A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}. Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only. A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction. This figure shows there is almost no difference between the simulations at all locations, rendering the tide-sure interaction negligible. In operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction. 
% I'm not sure if removing a background state is really appropriate here. There could be some differences in the background states because of the tides(ie where fresh water moves to, etc). I also think it is more appropriate to use the atmospheric pressure to remove the background state...

Finally, an assessment of the importance of including the tide-surge interaction is presented next. A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.  Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only. A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction. The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm lower than the maximum of the surge-only case.  In contrast, the tide-surge interaction at Victoria is hardly noticeable. Indeed, a spatial map of the modelled residual and surge-only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands. The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation. However, the effect is a relatively small underestimate of the surge elevation and no changes to the timing of the surge. So, in operational settings, performing a surge-only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed). A background state has not been removed from these plots. }
\label{fig:tidesurge}
\end{figure}
